---
title: "test"
author: "Xueyi Fan"
date: "June 20, 2016"
output: html_document
---

I choose the Income dataset which is contained by R.

Loading the data:

```{r}
library('arules')
library("arulesViz")
library('Matrix')










data(Income)
str(Income)


```


###generate a set of 50 or so (non-redundant) rules.

```{r}
#generate a set rules
rules.d <- apriori(Income)
rules.d
inspect(rules.d[1:5])

#remove the redundant rules
rules.d.sorted = sort(rules.d, by="lift")
subset.matrix <- is.subset(rules.d.sorted, rules.d.sorted)
subset.matrix[lower.tri(subset.matrix, diag=T)] <- NA 
redundant <- colSums(subset.matrix, na.rm=T) >=1
which(redundant)
rules.d.pruned <- rules.d.sorted[!redundant]
rules.d.pruned

```


### 1. Which rules make sense to you? Highlight the five best and five worst of your rule set. 
```{r}

inspect(rules.d.pruned[1:5])
inspect(rules.d.pruned[1600:1604])


```

The rules that if the person is crew and no survived, this person is a male makes sense to me. Because we know most of the crew in titanic are male and they sacrificed themselves and left survival opportunies to women and childern. 

The five best rule set are:  
{Class=Crew,Survived=No}       => {Sex=Male}  
{Class=Crew}                   => {Sex=Male}  
{Class=3rd,Sex=Male,Age=Adult} => {Survived=No}   
{Class=3rd,Sex=Male}           => {Survived=No}   
{Age=Adult,Survived=No}        => {Sex=Male}    

The five worst rule set are:
{Class=Crew}                   => {Age=Adult}  
{Class=1st}                    => {Age=Adult}  
{Survived=No}                  => {Age=Adult}    
{Sex=Male}                     => {Age=Adult}    
{}                             => {Age=Adult}    


### 2. How did you choose the level of support and confidence? 

```{r}

rules <- apriori(Income, parameter = list(support = 0.01, confidence = 0.99, minlen = 2))
rules
summary(rules)

```

According to the result of default rules, I want to focus on the rules with high confidence which is higher than 99% because I am not interest in a rule with low confidence. I want to generate enough rules that I can do my research, here, I choose a low support value 0.01.

Finally, I got 102772 rules.

### 3. What is the lift and conviction of your best and worst rules? 

Using the rules.d.pruned rules whidh contains a set of 11 rules.

```{r}

#best rule
interestMeasure(rules.d.pruned[1],measure= "lift")

interestMeasure(rules.d.pruned[1],transactions = Income,measure= "conviction")

#worst
interestMeasure(rules.d.pruned[11],measure= "lift")

interestMeasure(rules.d.pruned[11],transactions = Income,measure= "conviction")

```

The lift of best rules is 4.355387; the conviction of best rules is 7.437704;

The lift of worst rules is 4.088838; the conviction of worst rules is 4.92105.

### 4. Visualize your 50 association rules. Where do the best and worst end up in your plot? 


```{r}
plot(rules.d.pruned)

plot(rules.d.pruned, method = "graph", control = list(type="items"))

plot(rules.d.pruned, method = "paracoord", control = list(reorder = TRUE))

```

### 5. Does the model make sense?

This model can figure out the association between different items. 